### 2020-04-02 sections/acknowledgements.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section
Continues making changes suggested by KLM and CK

This project was partially funded by a Google Faculty Research Award to Karen Masters (\url{https://ai.google/research/outreach/faculty-research-awards/})

>>>>

This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P. Sloan Foundation. We would also like to thank the 2,340 volunteers who have submitted classifications to the \textit{Galaxy Builder} project, especially user EliabethB, whose presence on the \textit{Galaxy Builder} forum on top of a large number of galaxies modelled has been a huge help.

This project was partially funded by a Google Faculty Research Award to Karen Masters (\url{https://ai.google/research/outreach/faculty-research-awards/}), and Timothy Lingard acknowledges studentship funding from the Science and Technology Facilities Council (ST/N504245/1).

### 2020-04-02 sections/introduction.tex
Continues making changes suggested by KLM and CK

Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}).

@@

\subsection{Formation and Evolution of Spiral Arms}

@@

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, indicating that spirals exist for a long time or are continually rebuilt.

@@

Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations. Stronger bulges and more more massive central black holes are both linked to more tightly wound spiral arms, albeit with large amounts of scatter.

\subsection{Measuring galaxy pitch angle}

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}). One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.

A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by

\begin{equation}
r = A\,e^{\theta\tan\phi},
\end{equation}

where $\phi$ is the arm's pitch angle. One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}). Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

\begin{equation}
  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
\end{equation}

The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.

A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.

Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. \citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect. It is uncertain how variation between pitch angles of individual arms impacts this measurement.

>>>>

Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.

@@

@@

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, suggesting that spirals exist for a long time or are continually rebuilt.

@@

Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations. Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.

### 2020-04-02 sections/method.tex
Continues making changes suggested by KLM and CK

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies analysed in this paper are the 198 galaxies from \Lingard. These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.

Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the classifications of galaxies in this \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard. We remove any galaxies for which fewer than two spiral arms were identified \comment{why exclude one-armed?}. This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.

@@

\begin{equation}
r_\mathrm{arm} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

@@

% \begin{figure}
%   \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
%   \caption{The model used for galaxy pitch angle measurement for the sample.}
%   \label{fig:pymc3-model}
% \end{figure}

Our likelihood function is therefore

\begin{equation}
\mathcal{L} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

>>>>

% !TEX root = ../main.tex
\subsection{Measuring galaxy pitch angle}

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}). One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.

A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by

\begin{equation}
r = A\,e^{\theta\tan\phi},
\end{equation}

where $\phi$ is the arm's pitch angle. One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}). Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

\begin{equation}
  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
\end{equation}

The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.

A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.

Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. \citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect. It is uncertain how variation between pitch angles of individual arms impacts this measurement.

\subsection{The Galaxy Sample}
The galaxies analysed in this paper are the 198 galaxies from \Lingard. These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.

Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the classifications of galaxies in this \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard. We remove any galaxies for which fewer than two spiral arms were identified, as these do not provide our model with any information about inter-arm scatter. This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.

@@

\begin{equation}
r_\mathrm{arm} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

@@

The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution. Our likelihood function is therefore

\begin{equation}
\mathcal{L} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
\end{equation}

### 2020-04-02 sections/results.tex
Continues making changes suggested by KLM and CK

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies arm pitch angle with a high degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of 7.8\degree, which decreases to 6.7\degree and 5.8\degree for galaxies with three and four arms respectively.

Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.02 \pm 0.95$, confirming the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.

\subsection{Spiral Winding}
\label{section:spiral_winding}
In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We make use of the Kolmogorov-Smirnov test in a similar manner for comparison.

>>>>

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with a high degree of certainty (uncertainty less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle), however the pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$) is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.

Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.0\pm 1.0$, similar to the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.

\subsection{Spiral Winding}
\label{section:spiral_winding}
In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}. We also make use of the Kolmogorov-Smirnov test in a similar manner for comparison.
